---
title: 'Reproducible Research: Peer Assessment 1'
author: "Jizong Liang"
output: html_document
---
date: `r Sys.Date()`

```{r} global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
## setting the format for scientific notation, see detail from reference item 2
knitr::knit_hooks$set(inline = function(x) {
  knitr:::format_sci(x, 'md')
})

```

### Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./data/Dataset.zip",method="curl")

## Unzip the file folder downloaded into the "data" folder. 
unzip(zipfile="./data/Dataset.zip",exdir="./data")

ActivityData <- read.csv("./data/activity.csv")
```


### What is mean total number of steps taken per day?
```{r results="hide"}
stepsByDay <- tapply(ActivityData$steps, ActivityData$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r results='hide'}
(hist(stepsByDay,breaks=20,col="red", xlab="Total steps taken per day",
     ylab="Frenquency 20 bins",main="Total number of Steps taken per day"))
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMean=mean(stepsByDay)
stepsByDayMedian=median(stepsByDay)
```

* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

-----

### What is the average daily activity pattern?
```{r results="hide"}
(AveDailyActivityPattern <- aggregate(ActivityData$steps,
                                       list(ActivityData$interval), mean, na.rm=TRUE))
colnames(AveDailyActivityPattern)=c("Interval","meanSteps")
## see reference item 1 for further detail about aggregate()
```

##### 1. Make a time series plot
```{r}
(plot(AveDailyActivityPattern,type="l",xlab="5-minutes interval",
      ylab="average number of steps taken all days",
      main="Average daily activity pattern",cex.lab=0.7,cex.main=1))
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(AveDailyActivityPattern$meanSteps)
x=AveDailyActivityPattern$Interval[mostSteps]
time <- substr(as.POSIXct(sprintf("%04.0f", x), format='%H%M'), 12, 16)
```
* Most Steps at: `r time`

----

### Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
TotalNumMissingValues <- length(which(is.na(ActivityData$steps)))
```
* Number of missing values: `r TotalNumMissingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
Imputed <- ActivityData
Imputed$steps <- impute(ActivityData$steps, mean)
result <- length(which(is.na(Imputed$steps)))
```
* Number of missing values `r result`

##### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsByDay2 <- tapply(Imputed$steps, Imputed$date, sum, na.rm=TRUE)
hist(stepsByDay2, breaks=20, col="red",xlab="Total steps taken per day",
     ylab="Frenquency 20 bins", main="Total number of steps taken per day")
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
##### Do these values differ from the estimates from the first part of the assignment? What is the ##### impact of imputing missing data on the estimates of the total daily number of steps?

```{r results='asis' }
a=mean(stepsByDay2)
b=median(stepsByDay2)
```

* Mean (Imputed): `r a`      Mean: `r stepsByDayMean`
* Median (Imputed):  `r b`   Median:  `r stepsByDayMedian`
* From the result above, we observe that the Imputed Mean and Median are equal while the 
* The origianl Mean and Median are different. That indicates that the original distribution
* of the sample is skew. And the Imputed sample is corrected by removing the missing figure 
* and thus is approximately symmetric. 
 
----

### Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
Imputed$dateType <-  ifelse(as.POSIXlt(Imputed$date)$wday %in% c(0,6), 'weekend', 'weekday') 
## see Reference item 3 for further detail about date and time. 
```

##### 2. Make a panel plot containing a time series plot
```{r}
averagedImputed <- aggregate(steps ~ interval + dateType, data=Imputed, mean)
ggplot(averagedImputed, aes(interval, steps)) + 
    geom_line(col="steelblue") + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")+ 
    theme_bw(base_family = "Avenir", base_size = 10)
    
```

**Reference**

1:Using aggregate and apply in R, 05-22-2013
http://davetang.org/muse/2013/05/22/using-aggregate-and-apply-in-r/

2: Scientific notation in knitr: How to improve typography
http://stackoverflow.com/questions/26437835/scientific-notation-in-knitr-how-to-improve-typography

3: Handling date-times in R,Cole Beck, 08-30-2012
http://biostat.mc.vanderbilt.edu/wiki/pub/Main/ColeBeck/datestimes.pdf